---
title: "Bellabeat Case Study"
author: "Ashwin Jain"
date: "8/18/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing Packages

For this, I've installed these packages

'readxl', 'Rcpp', 'magrittr', 'dplyr'



```{r install}
library(readxl)
library(Rcpp)
library(magrittr)
library(dplyr)
```

# Importing Data

We first need to load in our data. It was downloaded as csv files which I then exported to Excel for visual purposes

```{r import}
dailyActivity_merged = read_excel("/Users/ashwinjain/Development/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.xlsx")
sleepDay_merged = read_excel("/Users/ashwinjain/Development/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.xlsx")

```

# Cleaning Data

## dailyActivity_merged

We will first work with the dailyActivity_merged dataset

```{r view dailyActivity_merged, echo=FALSE}

knitr::kable(dailyActivity_merged[1:5, ], caption="Preview of dailyActivities_merged")

col_before = colnames(dailyActivity_merged)
```



First, I deleted the SedentaryActiveDistance column - many of these entries had values over 0, which doesn't make sense because being sedentary and on the move are mutually exclusive


```{r delete SedentaryActiveDistance}
dailyActivity_merged <- select(dailyActivity_merged, -SedentaryActiveDistance)
```


List of columns before - 15 total
```{r before, echo=FALSE}
col_before
```

List of columns after - 14 total
```{r after, echo=FALSE}
colnames(dailyActivity_merged)
```


Second, I decided we should check for duplicated data and remove extras